---
title: "Portfolio Computational Musicology"
author: "Odile Grim"
date: "2021"
output: html_document
---

# Portfolio

For my corpus I would like to look at the repertoire of Bring Me The Horizon. Since the release of their first four albums, which can, according to Google, all be classified as "metalcore", their music has progressed more towards alternative-/pop-/electronic-rock and they even released an [EP with experimental and electronic music](https://open.spotify.com/album/6yC6aLzX5knhWpKDidwxft). Having been in their fanbase since around 2013, when *Sempiternal* was released, I myself experienced the changes in their music. Some people argue that only their "old" music is good and their new music is "too soft", leading to them not listening to the band's music anymore. My question is, what exactly has changed about their music, which aspects have remained similar to their early music and which haven't? Also, how has this affected their listeners/ fanbase? Did the demographics change? I definitely think that BMTH's music after 2013 has changed significantly compared to their earlier music and it will be most interesting to dive deeper into their work from that time on. As for specific tracks to look at *Pray for Plagues*, *Chelsea Smile* and *Shadow Moses* from their earlier work and *Drown*, *Mantra* and *Ludens* from their more recent work might be representative tracks, as they were all released as singles and thus viewed as important. Some of their more recent albums include different genres within them, so it might also be useful to examine those more closely and determine what and how big those differences are and whether they are significant.



```{r}
library(tidyverse)
library(spotifyr)
```

```{r}
old <- get_playlist_audio_features("", "6oeRMX8XZ8IS7MCDb6Z1BG")
new <- get_playlist_audio_features("", "4k98rVhcCw54dhLaUMiyLK")

oldvsnew <-
  bind_rows(
    old %>% mutate(category = "Old music"),
    new %>% mutate(category = "New music")
  )


ggplot(oldvsnew, aes(valence, energy, color = category, size = speechiness)) + 
  geom_point(alpha = 0.5) + 
  theme(legend.position = "none") +
  geom_label(
    aes(label = "Old music"),
    x = 0.1, y = 0.68,
    label.padding = unit(0.3, "lines"),
    label.size = 0.4,
    color = "black",
    fill = "lightblue"
  )+
  geom_label(
    aes(label = "New music"),
    x = 0.6, y = 0.99,
    label.padding = unit(0.3, "lines"),
    label.size = 0.4,
    color = "black",
    fill = "pink"
  )+
  labs(title = "Bring Me The Horizon's old vs. new music")

```

# Bring Me The Horizon's old vs. new music

For this data I compared the older music of BMTH to their newer music. For their old music I simply took their first two albums *Count Your Blessings* (2006) and *Suicide Season* (2008) and for the new music I took their most recent release, the EP *POST HUMAN: SURVIVAL HORROR* (2020) and their most recent album *amo* (2019). The reason that I did not simply use their two most recent releases for the "new music" category, is that I do not think that it would be relevant. If I would use their second or third most recent work instead of *amo*, I would be using a live album (*Live at the Royal Albert Hall* (2020)) or an album that was more of an experiment and not necessarily a representation of their music style (*Music to listen to~* (2019)), although that would also be an interesting thing to look at.

Looking at the plot that I have made with the data, you can easily see that the overall mood of their old vs. new music is significantly different. Taking into account that high energy and low valence means 'angry music' and high valence and high energy means 'happy music', you could conclude that their new music is overall much happier than their old music, which isn't happy at all. 

**(((to be continued)))**
